Introducing Distributed Dynamic Data-intensive (D3) Science: Understanding Applications and Infrastructure

### 2015-09-13 3dpas_intro.tex
further fixes in introduction.

The landscape of scientific and enterprise computing is being fundamentally
altered by the prodigious volumes of data.  % Not only are the volumes of data
% produced by scientific applications increasing rapidly, there is also a
% corresponding increase in the amount of data that needs to be processed and
% analyzed.
With the increase in data volumes, types, and sources of data, an
important concomitant development has been an increase in the requirement for
and complexity of the infrastructure needed to support these requirements.
Furthermore, data are often naturally distributed and have to be processed in
situ; occasionally data that are localized have to be distributed in order to be
processed, either for reasons of performance or for collaborative analytics.
The scale and significance of varying data source/destination and
transfer/production rates makes temporal variations important, which in turn
eventually impact the design and development of infrastructure to support them:
resource scheduling, data placement, and transfer decisions that were statically
determinable at small scales, emerge as dynamic and important considerations at
large scales of data and distribution.

@@

The challenges of distributed data-intensive applications span beyond the
management and curation of large volumes of data. A critical requirement is to
make sense of this data, i.e., transform the data into information en route to
assembling knowledge. Among other things, the transformation from data to
information requires extensive amounts of computational processing of the data;
thus the data-intensive nature of the data-intensive science problem is strongly
coupled with compute-intensive requirements.

@@

Subjecting a range of otherwise distinct applications to a
common analytical framework provides advantages to the infrastructure
scientists and cyberinfrastructure experts, namely suggestions for
where effort is both needed and likely to yield maximum returns. In
other words, as D3 applications evolve, our analysis can provide
guidance on where to focus effort on infrastructure.  Last but not
least, the advantages of providing a common but extensible vocabulary
that can be used by the entire community, application end-users,
developers and infrastructure scientists alike, in an emerging field
cannot be overstated. One may argue that such vocabulary and
terminology has been missing from the practice of distributed
cyberinfrastructure in general, with less than desirable consequences.

@@

The focus of this paper is on a subspace of the large set of problems associated
with data-intensive sciences, namely that which is characterized by {\it both}
distributed and dynamic aspects. We introduce the concept of Dynamic Distributed
Data-intensive science and refer to this subspace of data-intensive science as
D3 science.  Analyzing the applications along an analytical space spanned by the
three ``D''s of data-intensive, distributed, and dynamic provides a general and
effective framework.  The diverse set of requirements to support D3 scenarios
points to the need for a careful understanding of D3 characteristics as well as
sophisticated, extensible, and skillfully architected infrastructure and
programming systems.

>>>>

The landscape of scientific and enterprise computing is being fundamentally
altered by the prodigious volumes of data.  % Not only are the volumes of data
% produced by scientific applications increasing rapidly, there is also a
% corresponding increase in the amount of data that needs to be processed and
% analyzed.
With the increase in data volumes, types, and sources of data, a concomitant
development has been an increase in the sophistication and complexity of the
infrastructure needed to support these requirements. For example, data are often
naturally distributed and have to be processed in situ; occasionally data that
are localized have to be distributed in order to be processed, either for
reasons of performance or for collaborative analytics.  The scale and
significance of varying data source/destination and transfer/production rates
makes temporal variations important, which in turn eventually impact the design
and development of infrastructure to support them: resource scheduling, data
placement, and transfer decisions that were statically determinable at small
scales, emerge as dynamic and important considerations at large scales of data
and distribution.

@@

The challenges of distributed data-intensive applications span beyond the
management and curation of large volumes of data. A critical requirement is to
transform the data into information en route to assembling knowledge. Among
other things, the transformation from data to information requires extensive
amounts of computational processing of the data; thus the data-intensive nature
of the data-intensive science problem is strongly coupled with compute-intensive
requirements.

@@

The focus of this paper is on a subspace of the large set of problems associated
with data-intensive sciences, namely that which is characterized by {\it both}
distributed and dynamic aspects. We introduce the concept of Dynamic Distributed
Data-intensive science and refer to this subspace of data-intensive science as
D3 science.

Analyzing the applications along an analytical space spanned by the three ``D''s
of data-intensive, distributed, and dynamic provides a general and effective
framework.  Subjecting a range of otherwise distinct applications to a common
analytical framework provides advantages to the infrastructure scientists and
cyberinfrastructure experts, namely suggestions for where effort is both needed
and likely to yield maximum returns.  The diverse set of requirements to support
D3 scenarios points to the need for a careful understanding of D3
characteristics as well as sophisticated, extensible, and skillfully architected
infrastructure.

As D3 applications evolve, our analysis can provide guidance on where to focus
effort on infrastructure.  Last but not least, the advantages of providing a
common but extensible vocabulary that can be used by the entire community,
application end-users, developers and infrastructure scientists alike, in an
emerging field cannot be overstated. One may argue that such vocabulary and
terminology has been missing from the practice of distributed
cyberinfrastructure in general, with less than desirable consequences.

@@
